Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/System.tex
merging Nirav's changes to system (using word spontaneous)
maybe fixed another of adamc's comments?

We have dedicated channels between each parent-child pair. The three channels
are indexed by the name of the child that uses that channel (denoted by
$\textit{ch}[c]$ for the \textit{ch} channel associated
with cache $c$).  Figure~\ref{format} shows the message formats
transmitted in the 3
channels between caches. In each of the channels $ch$, a message $m$
can be enqueued ($\enq(ch, m)$) or dequeued ($\deq(ch)$), and the
channels can be examined for the first element ($\first(ch)$) or for
availability of any message in the channel
($\avail(ch)$). Each of these channels start out empty, \ie{} $\forall ch. \;
\neg \avail(ch)$, initially.

\begin{figure}
\begin{subfigure}{6.8cm}
\begin{tabular}{|lp{5.8cm}|}
\hline
\multicolumn{2}{|c|}{\Reqcp}\\
\hline
\from: & Current state of child\\
\myto: & Desired (higher) state of child\\
\addr: & Address associated with the request\\
\hline
\hline
\multicolumn{2}{|c|}{\Respcp}\\
\hline
\from: & Current state of child\\
\myto: & Downgraded state of child\\
\addr: & Address associated with the request\\
\data: & Data associated with the address, if necessary\\
\hline
\end{tabular}
\end{subfigure}
\begin{subfigure}{5.4cm}
\begin{tabular}{|lp{4.4cm}|}
\hline
\multicolumn{2}{|c|}{\Mpc}\\
\hline
\typ: & \Req{} (for requests) and \Resp{} (for responses)\\
\myto: & Desired (lower) directory state of a child for \Req{} and upgraded
directory state of a child for \Resp{}\\
\addr: & Address associated with the request\\
\data: & Data associated with the address if necessary (for \Resp{} only)\\
\hline
\end{tabular}
\end{subfigure}
\caption{Data types for messages between caches
\adamc{Names here don't match those in the preceding text.}}
\label{format}
\end{figure}

@@

We first give an intuition on how the transitions work in the common case.
A cache can decide, unprovoked, to upgrade its state,
in which case it sends an upgrade request to its parent. The parent then makes
a local decision on whether to send a response to the requesting child or not
(based on its directory approximation, and its own state). If its own state is
lower the requested upgrade, then it cannot handle the request. The parent
must instead decide to upgrade its state -- the transition for doing so has
nothing to do with the incoming request.  Once the parent's state is not lower
than the requested upgrade, it makes sure that the rest of its children are
``compatible'' with the requested upgrade. If not, the parent must send
requests to the incompatible children to downgrade -- again the transition for
doing so has nothing to do with the incoming request. If the children are
compatible and its own state is higher, then the request can be handled and a
response is sent. Each of the transitions is explained in greater detail below.

>>>>

We have dedicated channels between each parent-child pair. The three channels
are indexed by the name of the child that uses that channel (denoted by
$\textit{ch}[c]$ for the \textit{ch} channel associated
with cache $c$).  Figure~\ref{format} shows the fields of the messages
transmitted in the 3 channels between caches. In each of the channels $ch$, a
message $m$ can be enqueued ($\enq(ch, m)$) or dequeued ($\deq(ch)$), and the
channels can be examined for the first element ($\first(ch)$) or for
availability of any message in the channel ($\avail(ch)$). Each of these
channels start out empty, \ie{} $\forall ch. \; \neg \avail(ch)$, initially.

\begin{figure}
\begin{subfigure}{6.8cm}
\begin{tabular}{|lp{5.8cm}|}
\hline
\multicolumn{2}{|c|}{A message in \cpReq}\\
\hline
\from: & Current state of child\\
\myto: & Desired (higher) state of child\\
\addr: & Address associated with the request\\
\hline
\hline
\multicolumn{2}{|c|}{A message in \cpResp}\\
\hline
\from: & Current state of child\\
\myto: & Downgraded state of child\\
\addr: & Address associated with the request\\
\data: & Data associated with the address, if necessary\\
\hline
\end{tabular}
\end{subfigure}
\begin{subfigure}{5.4cm}
\begin{tabular}{|lp{4.4cm}|}
\hline
\multicolumn{2}{|c|}{A message in \pc}\\
\hline
\typ: & \Req{} (for requests) and \Resp{} (for responses)\\
\myto: & Desired (lower) directory state of a child for \Req{} and upgraded
directory state of a child for \Resp{}\\
\addr: & Address associated with the request\\
\data: & Data associated with the address if necessary (for \Resp{} only)\\
\hline
\end{tabular}
\end{subfigure}
\caption{Data types for messages between caches
\adamc{Names here don't match those in the preceding text.}}
\label{format}
\end{figure}

@@

We first give an intuition on how the transitions work in the common case.
A cache can decide, spontaneously, to upgrade its state,
in which case it sends an upgrade request to its parent. The parent then makes
a local decision on whether to send a response to the requesting child or not
(based on its directory approximation, and its own state). If its own state is
lower the requested upgrade, then it cannot handle the request. The parent
must instead decide to upgrade its state -- the transition for doing so has
nothing to do with the incoming request.  Once the parent's state is not lower
than the requested upgrade, it makes sure that the rest of its children are
``compatible'' with the requested upgrade. If not, the parent must send
requests to the incompatible children to downgrade -- again the transition for
doing so has nothing to do with the incoming request. If the children are
compatible and its own state is higher, then the request can be handled and a
response is sent. Each of the transitions is explained in greater detail below.
